Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2019-09-25 manuscript/manuscript.tex
minor ms changes

\clearpage
\section{Results}

@@

\clearpage

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to epistatic relationships
among alleles altered by frequency changes or recombination breaking up
selective alleles.  Such reversals of selective dynamics could be occurring at
other timepoints but the signal of a change in the direction of selection at
particular loci may be washed out when we calculate our genome-wide average
temporal covariances. To address this limitation, we calculated the
distribution of the temporal covariances over 100kb genomic windows \vb{which
we refer to as \emph{windowed covariances}} (Figure \ref{fig:figure-3}
\vb{shows these distributions} pooling across all replicates; see Supplementary
Figure \ref{suppfig:barghi-offset-replicate-panels} for individuals
replicates). The covariance of each tile will be noisy, due to sampling and
genetic drift, and the neutral distribution of the covariance is complicated
due to linkage disequilibria (which can occur over long physical distances in
E\&R and selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To
address this, we have developed a permutation-based procedure that constructs a
null distribution by randomly flipping the signs of the  allele frequency
changes per-genomic window. This destroys the systematic covariances created by
linked selection and creates a sampling distribution of the covariances
spuriously created by neutral genetic drift while preserving the complex
dependencies between adjacent loci created by linkage disequilibrium. This
empirical neutral null distribution is conservative in the sense that the
variances of the covariances are wider than expected under drift alone as they
include the effect of selection on the allele frequency change within a
time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
striking pattern that the windowed covariances not only decay towards zero, but
in fact become negative through time, consistent with many regions in the
genome having had a reversed fitness effect at later timepoints.

@@

@@

  \caption{\vb{\textbf{A}: Histogram of original \textcite{Bergland2014-ij}
    seasonal p-values and p-values creating by randomly permuting the seasons
    at each locus. \textbf{B}: Histogram of original \textcite{Bergland2014-ij}
    p-values alongside all unique permutations (ignoring symmetries that lead
    to identical p-values).}}

@@

  \caption{Temporal covariances from the \textcite{Bergland2014-ij} study, from
    varying reference generations (e.g. rows along the temporal covariance
    matrix). Each covariance is labeled indicating whether the covariance is
    between two like seasonal transitions (e.g. the covariance between allele
    frequency changes from fall to spring in one year, and fall to spring in
    another or two dislike seasons (e.g. the covariance between fall to spring
    in one year, and spring to fall in another year). Covariances between like
    transitions are expected to be positive when there is a genome-wide effect
    of fluctuating selection (and these labels are colored blue), while
    covariances between dislike transitions are expected to be negative (and
    these labels are colored red). 95\% confidence intervals were constructed
    by a block-bootstrapping procedure where the blocks are megabase tiles.}

>>>>

\section{Results}

@@

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to epistatic relationships
among alleles altered by frequency changes or recombination breaking up
selective alleles.  Such reversals of selective dynamics could be occurring at
other timepoints but the signal of a change in the direction of selection at
particular loci may be washed out when we calculate our genome-wide average
temporal covariances. To address this limitation, we calculated the
distribution of the temporal covariances over 100kb genomic windows which we
refer to as \emph{windowed covariances} (Figure \ref{fig:figure-3} shows these
distributions pooling across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance of each tile will be noisy, due to sampling and genetic drift, and
the neutral distribution of the covariance is complicated due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
this, we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the  allele frequency changes
per-genomic window. This destroys the systematic covariances created by linked
selection and creates a sampling distribution of the covariances spuriously
created by neutral genetic drift while preserving the complex dependencies
between adjacent loci created by linkage disequilibrium. This empirical neutral
null distribution is conservative in the sense that the variances of the
covariances are wider than expected under drift alone as they include the
effect of selection on the allele frequency change within a time-interval, just
not between time-intervals. We see (Figure \ref{fig:figure-3} A and B) that
windowed temporal covariances between close timepoints are skewed positive (a
heavy right tail), while between more distant timepoints these windowed
temporal covariances tend to shift to become more negative (a heavy left tail).
We quantified the degree to which the left and right tails are inflated
compared to the null distribution as a function of time, and see excesses in
both tails in Figure \ref{fig:figure-3} C. This finding is also robust to
sign-permuting allele frequency changes on a chromosome-level, the longest
extent that gametic linkage disequilibria can extend (Supplementary Figure
\ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that the
windowed covariances not only decay towards zero, but in fact become negative
through time, consistent with many regions in the genome having had a reversed
fitness effect at later timepoints.

@@

  \caption{Temporal covariances from the \textcite{Bergland2014-ij} study, from
    varying reference generations (e.g. rows along the temporal covariance
    matrix). Each covariance is labeled indicating whether the covariance is
    between two like seasonal transitions (e.g. the covariance between allele
    frequency changes from fall to spring in one year, and fall to spring in
    another) or two dislike seasons (e.g. the covariance between fall to spring
    in one year, and spring to fall in another year). Covariances between like
    transitions are expected to be positive when there is a genome-wide effect
    of fluctuating selection (and these labels are colored blue), while
    covariances between dislike transitions are expected to be negative (and
    these labels are colored red). 95\% confidence intervals were constructed
    by a block-bootstrapping procedure where the blocks are megabase tiles.}

@@

  \caption{\textbf{A}: Histogram of original \textcite{Bergland2014-ij}
    seasonal p-values and p-values creating by randomly permuting the seasons
    at each locus. \textbf{B}: Histogram of original \textcite{Bergland2014-ij}
    p-values alongside all unique permutations (ignoring symmetries that lead
    to identical p-values).}

@@
